Avoiding tipping points in fisheries management: A Bayesian Non-Parametric approach

### 2014-06-23 manuscripts/manuscript.Rmd
steve edits

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Ecological dynamics are frequently complex and difficult to measure,
making uncertainty in our understanding a persistent challenge.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (but stretching even further back in time), marked by  the
seminal contributions of @Clark1976, @Clark2009 and @Walters1978,
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of
understanding how to management human intervention into natural systems.
Concomittantly, there has been increasing recognition of the importance of
Allee population sizes or `tipping points' [@Scheffer2001;
@Polasky2011] that management actions may inadvertently
cause populations to cross.  @Scheffer2009 emphasized the difficulties of
formulating even qualitatively correct models of the underlying processes.

@@

Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
'best-case' scenario for the parametric model performance relative to
the GP.   This is abest-case scenario for the parametric models because the underlying dynamics are always simuated
from one the parametric models in the set, whereas in reality we never
have the "true" model among our candidates.  By choosing one-dimensional
models with few parameters, we limit the chance that poor performance
will be due to our inability to estimate parameters well, something
that becomes a more severe problem for higher-dimensional parametric
models. Further, the parametric models we consider are those most commonly
used in modeling stock-recruitment dynamics or to model sudden transitions
between alternative stable states.

@@

We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are. NOTE: EQUATION NUMBERS ARE ALL MESSED UP. THIS SHOULD BE EQ 1, BUT DOES NOT APPEAR ON THE PDF YOU SENT TO ME

$$ X_{t+1}= Z_t f(X_t) $$

@@

To construct the value function, we consider a return when $X_t=x$ and harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We assume that future harvests are discounted relative to current ones at a constant rate of discount $\delta$ and ask for the harvest policy that maximizes total discounted harvest between the current time $t$ and a final time $T$, that is we seek to maximize over choices of harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$, where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the expectation over the process stochasticity in the future population state.

In order to find that policy, we introduce the value function$V_t(x_t)$ representing the total discounted catch from time $t$ onwards given that $X_t=x$.  This value function satisfies an equation of SDP  [ @Mangel1988; @Clark2000; @Clark2009], sometimes called the Bellman-Hamilton-Jacobi equation [@Mangel2014],

\begin{equation}
V_t(X_t) = \max_{h_t} \lbrace R(h_t, X_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] \rbrace
\end{equation}

where expectation is taken over all possible values of the next state, $X_t{+1}$, and maximized over all possible choices of harvest, $h_t$.  That is, at time $t$, when population size is $x$ and harvest $h_t$ is applied, the immediate return is $R(h_t, X_t)$. In the classical case, the sole source of uncertainty is the process stochasticity term, $Z$, and thus the expectation above is equivalent to taking expectations over $Z$. That is

$$ \mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] $$

@@

In a Bayesian decision framework, the parameters governing the dynamics are also uncertain, so the
expectation in (1) involves averaging over the posterior distribution for the parameters, as well. That
is,

$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$

@@

$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta,\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | X_t, h_t \right] \}$$

We consider the finite time problem with $T=$ `r MaxT`, which we solve using value iteration algorithm [@Mangel1988; @Clark2000].

@@

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
model [@Myers1995], equations (@ricker)-(@myers), which we parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate. In the Ricker model,

(@ricker) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } $$

@@

(@allen) $$ f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} $$

@@

(@myers) $$ f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}}$$

@@

We consider a period of `r Tobs` years of training data: long
enough that the estimates do not depend on the particular realization,
while longer times are not likely to provide substantial improvement
(i.e. the results are not sensitive to this interval).  Each of the models
(described below) is fit to the same training data (Figure 1).

@@

We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
[@RTeam] using jags, [@R2jags]).  We choose uniform priors
for all parameters (See appendix Tables S1-S3; R code provided). We
show one-step-ahead predictions of these model fits in Figure 1.
We tested each chain for Gelman-Rubin convergence and results were
robust to longer runs.  For each simulation we also applied several
commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
to identify the best fitting model.

@@

The use GP methods to formulate a predictive
model is relatively new in the context of modeling dynamical systems
[@Kocijan2005], and was first introduced in the context ecological
modeling and fisheries management in @Munch2005.  GP models have
subsequently been used to test for the presence of Allee effects
[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
determine temporal variation in food availability [@Sigourney2012], and
provide a basis for identifying model-misspecification [@Thorson2014].
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

A GP is an infinite dimensional (because it is a function of a continuous input
variable) stochastic process for which any realization is a multivariate
normal distribution. Thus, to characterize the GP we need a mean function
and a covariance matrix. We proceed as follows.

@@

We use vague inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form

@@

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the two parameters of the GP (Figure
S13, code in appendix), under weakly informative Gaussian priors
(see parameters in table S5). Since the posterior distributions differ
substantially from the priors (Figure S13), we can be assured that most
of the information in the posterior comes from the data rather than the
prior belief.

<!-- SDP via the model  FIXME much more detail here-->

@@

To ensure our results are robust to our choice of parameters, 
consider 96 different scenarios, described in detail below.   To begin, to help  better
understand the process, we first describe in detail the
results of a single scenario.

@@

NOTE: IN THE PDF YOU SENT THERE IS WEIRD STUFF ON LINE 237

```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))
xtable::xtable(rbind(dictable, aictable, bictable))
```

```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
xtable::xtable(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model. As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion. This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.")
```

@@

NOTE: THERE IS NO MENTION OF FIGURE 3 IN THE TEXT

@@

In Figure 5, we show the consequences of managing 100 replicate realizations of the simulated
fishery under each of the policies estimated..
As anticipated from the policy curves, the structurally correct model
under-harvests, leaving the stock to vary around its unfished optimum.
The structurally incorrect Ricker model over-harvests the population
passed the tipping point consistently, resulting in the immediate crash
of the stock and thus derives minimal long-term catch.

@@

These results are not sensitive to the modeling details of the simulation.
The GPDP estimate remains very close to the optimal solution obtained
by knowing the true model across changes to the training simulation,
noise scale, parameters or structure of the underlying model. We consider
both a Latin hypercube approach and a more systematic investigation of the
parameters most responsible for impacting the transition rate: the relative 
distance to the Allee threshold and the noise effects.

@@

(see @Gardiner2009 or @Mangel2006 for the derivation).
As a result, the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be modeled
near-optimally. Models that underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

@@

Model selection criteria trade off model complexity and fit to the data. 
When the data come from a limited region of state-space -- as is necessarily 
the case whenever there is a potential concern about tipping point dynamics --
simpler models therefore fit just as well and tend to be selected more foten than complex 
ones.  This approach would be appropriate when the dynamics can be expected 
to remain in the region of the training data; for instance, if we only 
considered the forecasting accuracy of the unfished population dynamics under
each model.

@@

### GPDP population dynamcis capture larger uncertainty in regions where the data are poor

<!-- FIXME More emphasis/explanation on _how_ it does this -->

The parametric models perform worst when they propose a management strategy
outside the range of the observed data. The non-parametric Bayesian approach, 
in contrast, allows a predictive model that expresses a great deal of uncertainty
about the probable dynamics outside the observed range, while retaining very
good predictive accuracy in the range observed.  The management policy 
dictated by the GP balance this uncertainty against
the immediate value of the harvest, and act to stabilize the population 
dynamics in a region of state space in which the predictions can be 
reliably reflected by the data.

@@

Lastly, it should be noted that outside the data, the GP reverts to
the prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the
SDP.  In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its parameters
$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
In practice, both the choice of mean and the choice of the covariance
function may be chosen to confer particular biological properties,
as well as more biologically informed priors for $\ell$ and $\sigma$.
In principle, this may allow a manager to improve the performance of the
GPDP by adding only enough additional detail as is justified.
For instance, it would be possible to use a linear or a Ricker-shaped
mean in the prior without making the much stronger assumption that the
Ricker is the structurally correct model [@Sugeno2013a]. Future research
should focus on identifying criteria to ensure the prior and the value function
are chosen appropriately for the problem at hand.

@@

NOTE: THE PDF HAS TWO }} IN LINE 351

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}}
\setcounter{figure}{0}}
\setcounter{table}{0}}

@@

setA = c(r = p1[1], K = p2[1], theta = p3[1])
setB = c(r = p1[2], K = p2[2], theta = p3[2])
setC = c(r = p1[3], K = p2[3], C = p3[3])
setD = c(r = p1[4], K = p2[4], C = p3[4])
AllenParameterSets <- rbind(setA, setB)
MyersParameterSets <- rbind(setC, setD)

>>>>

Decision making under uncertainty is a ubiquitous challenge in the
management of human intervention in natural resources and conservation.
Ecological dynamics are frequently complex and difficult to measure,
making uncertainty in our understanding a persistent challenge.
Decision-theoretic approaches provide a framework to determine the
best sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified [@Fischer2009].  Over the last
four decades (but stretching even further back in time), marked by  the
seminal contributions of @Clark1976, @Clark2009 and @Walters1978,
dynamic optimization methods, particularly Stochastic Dynamic Programming
(SDP), have become increasingly important as a means of
understanding how to manage human intervention into natural systems.
Simultaneously, there has been increasing recognition of the importance of
Allee population sizes or `tipping points' [@Scheffer2001;
@Polasky2011] that management actions may inadvertently
cause populations to cross.  @Scheffer2009 emphasized the difficulties of
formulating even qualitatively correct models of the underlying processes.

@@

Dynamic optimization  requires characterizing the dynamics of a state variable (or variables), a control action, and a value function.
In this paper, we consider only a single state variable, so that the ideas can be described as simply as possible. That is, we focus on a class of one-dimensional models of population dynamics as a
'best-case' scenario for the parametric model performance relative to
the GP.   This is a best-case scenario for the parametric models because the underlying dynamics are always simuated
from one of the parametric models in the set, whereas in reality we never
have the "true" model among our candidates.  By choosing one-dimensional
models with few parameters, we limit the chance that poor performance
will be due to our inability to estimate parameters well, something
that becomes a more severe problem for higher-dimensional parametric
models. Further, the parametric models we consider are those most commonly
used in modeling stock-recruitment dynamics or to model sudden transitions
between alternative stable states.

@@

We let $X_t$ denote the size (numbers or biomass) of the focal population at time $t$ and assume that in the absence of take its dynamics are.

\begin{equation}
X_{t+1}= Z_t f(X_t) \label{eq1} 
\end{equation}

@@

To construct the value function, we consider a return when $X_t=x$ and
harvest $h_t=h$ denoted as the reward, $R(x,h)$. For example, if the
return is simply the harvest at time $t$, then $R(x,h)=min(x,h)$. We
assume that future harvests are discounted relative to current ones at
a constant rate of discount $\delta$ and ask for the harvest policy
that maximizes total discounted harvest between the current time $t$
and a final time $T$. That is, we seek to maximize over choices of
harvest $\mathbf{E}_{X_t+1} [ \sum_{t = 0}^{T}  R_t(X_t, h_t) \delta^t]$,
where the state dynamics are given by Eqn 1 and $\mathbf{E}$ denotes the
expectation over the process stochasticity in the future population state.

In order to find that policy, we introduce the value function $V_t(x_t)$
representing the total discounted catch from time $t$ onwards
given that $X_t=x_t$.  This value function satisfies an equation of
SDP  [@Mangel1988; @Clark2000; @Clark2009], sometimes called the
Hamilton-Jacobi-Bellman equation [@Mangel2014],

\begin{equation}
V_t(x_t) = \max_{h_t} \lbrace R(h_t, x_t) + \delta \cdot \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] \rbrace
\end{equation}

where expectation is taken over all possible values of the next state,
$X_{t+1}$, and maximized over all possible choices of harvest, $h_t$.
That is, at time $t$, when population size is $x_t$ and harvest $h_t$ is
applied, the immediate return is $R(h_t, x_t)$. In the classical case,
the sole source of uncertainty is the process stochasticity term, $Z$,
and thus the expectation above is equivalent to taking expectations over
$Z$. That is

$$ \mathbf{\mathrm{E}}_{X_{t+1}} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{Z} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] $$

@@

In a Bayesian decision framework, the parameters governing the dynamics
are also uncertain, so the expectation involves averaging over the
posterior distribution for the parameters, as well. That is,

$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | x_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(x_t - h_t))  | x_t, h_t \right] \}$$

@@

$$\mathbf{\mathrm{E}}_{X_t+1} \left[ V_{t+1}( X_{t+1}) | X_t, h_t \right] = \mathbf{\mathrm{E}}_{\theta|\mathrm{data}} \{ \mathbf{\mathrm{E}}_{f, Z | \theta, \mathrm{data}} \left[ V_{t+1}( Z f(X_t - h_t))  | x_t, h_t \right] \}$$

We consider the finite time problem with $T=$ `r MaxT`, which we
solve using the standard value iteration algorithm [see @Mangel1988;
@Clark2000].

@@

We consider three candidate parametric models of the stock-recruitment
dynamics: The Ricker model, the Allen model [@Allen2005], the Myers
model [@Myers1995], Eqns \eqref{ricker}-\eqref{myers}, which we parameterize as follows for $f(S_t,p)$ in Eqn 1. In all three, we let $K$ denote the carrying capacity and $r$ the maximum per capita growth rate. In the Ricker model,

\begin{equation}
f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K} \right) } \label{ricker}
\end{equation}

@@

\begin{equation}
f(S(t)|p) = S_t e^{r \left(1 - \frac{S_t}{K}\right)\left(S_t - X_c\right)} \label{allen}
\end{equation}

@@

\begin{equation}
f(S(t) | p)  = \frac{r S_t^{\theta}}{1 + \frac{S_t^\theta}{K}} \label{myers}
\end{equation}

@@

We consider a period of `r Tobs` years of training data: long
enough that the estimates do not depend on the particular realization,
while longer times are not likely to provide substantial improvement.  
Each of the models is fit to the same training data (Figure 1).

@@

We inferred posterior distributions for the parameters of each model by Gibbs sampling (@Gelman2003 implemented in R
[@RTeam] using `jags`, [@R2jags]).  We choose uniform priors
for all parameters (See appendix Tables S1-S3; R code provided). We
show one-step-ahead predictions of these model fits in Figure 1.
We tested each chain for Gelman-Rubin convergence and results were
robust to longer runs.  For each simulation we also applied several
commonly used model selection criteria (AIC, BIC, DIC, see @Burnham2002)
to identify the best fitting model.

@@

The use of GP methods to formulate a predictive
model is relatively new in the context of modeling dynamical systems
[@Kocijan2005], and was first introduced in the context ecological
modeling and fisheries management in @Munch2005.  GP models have
subsequently been used to test for the presence of Allee effects
[@Sugeno2013a], estimate the maximum reproductive rate [@Sugeno2013b],
determine temporal variation in food availability [@Sigourney2012], and
provide a basis for identifying model-misspecification [@Thorson2014].
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

A GP is an infinite dimensional (because it is a function of a continuous input
variable) stochastic process for a finite sample of points from any realization is a multivariate
normal distribution. Thus, to characterize the GP we need a mean function
and a covariance function. We proceed as follows.

@@

We use inverse Gamma priors on both the lengthscale $\ell$ and $\sigma$, of the form

@@

We use a Metropolis-Hastings Markov Chain Monte Carlo (@Gelman2003) to
infer posterior distributions of the two parameters of the GP (Figure
S13, code in appendix). Since the posterior distributions differ
substantially from the priors (Figure S13), we can be assured that most
of the information in the posterior comes from the data rather than the
prior belief.

@@

To ensure our results are robust to our choice of parameters, 
consider 96 different scenarios, described in detail below.   To help  better
understand the process, we first describe in detail the
results of a single scenario.

@@

```{r deviances, dependson=c("dic_calc"), include=FALSE, echo=FALSE}
allen_deviance  <- - posterior.mode(pardist[,'deviance'])
ricker_deviance <- - posterior.mode(ricker_pardist[,'deviance'])
myers_deviance  <- - posterior.mode(myers_pardist[,'deviance'])
true_deviance   <- 2*estf(c(p, sigma_g))
mle_deviance    <- 2*estf(c(est$p, est$sigma_g))
aictable <- data.frame(Allen = allen_deviance + 2*(1+length(bayes_pars)),  # +1 for noise parameter
                       Ricker = ricker_deviance + 2*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + 2*(1+length(myers_bayes_pars)), 
                       row.names = c("AIC"))
bictable <- data.frame(Allen = allen_deviance + log(length(x))*(1+length(bayes_pars)), 
                       Ricker = ricker_deviance + log(length(x))*(1+length(ricker_bayes_pars)),
                       Myers = myers_deviance + log(length(x))*(1+length(myers_bayes_pars)), 
                       row.names = c("BIC"))
pandoc.table(rbind(dictable, aictable, bictable))
```

```{r Table1, dependson="deviances", include=TRUE, results="asis", echo=FALSE, cache=FALSE}
pandoc.table(rbind(dictable, aictable, bictable), caption="Model selection scores for several common criteria all (wrongly) select the simplest model. As the true (Allen) model is not distinguishable from the simpler (Ricker) model in the region of the observed data, this error cannot be avoided regardless of the model choice criterion. This highlights the danger of model choice when the selected model will be used outside of the observed range of the data.")
```

@@

Figure 3 illustrates the performance of the GP and parametric models
outside the observed training data. The mean trajectory under the
underlying model is shown by the black dots, while the corresponding
prediction made by the model shown by the box and whiskers plots.
Predictions are based on the true expected value in the previous time
step.  Predicted distributions that lie entirely above the expected
dynamics indicate the expectation of stock sizes higher than what is
actually expected. The models differ both in their expectations and
their uncertainty (colored bands show two standard deviations away).
Note that the GP is particularly uncertain about the dynamics relative
to structurally incorrect models like the Ricker.

@@

In Figure 5, we show the consequences of managing 100 replicate
realizations of the simulated fishery under each of the policies
estimated.  As anticipated from the policy curves, the structurally
correct model under-harvests, leaving the stock to vary around its
unfished optimum.  The structurally incorrect Ricker model over-harvests
the population passed the tipping point consistently, resulting in the
immediate crash of the stock and thus derives minimal long-term catch.

@@

These results are not sensitive to the modeling details of the simulation.
The GPDP estimate remains very close to the optimal solution obtained
by knowing the true model across changes to the training simulation,
noise scale, parameters or structure of the underlying model. We consider
both a Latin hypercube approach and a more focused investigation of the
parameters most responsible for impacting the transition rate: the relative 
distance to the Allee threshold and the noise effects.

@@

(see @Gardiner2009 or @Mangel2006 for the derivation).
As a result, the impact of using a model which underestimates the risk of harvesting
past the critical point is much more severe, since this such a situation
occurs more often.  Conversely, with large enough distance between the
optimal escapement and unstable points relative to the noise, the chance
of a transition becomes vanishingly small and all models can be estimated 
near-optimally. Models that underestimate the cost incurred by stock
sizes fluctuating significantly below the optimal escapement level will
not perform poorly as long as those fluctuations are sufficently rare.

@@

Model selection criteria trade off model complexity and fit to the data. 
When the data come from a limited region of state-space -- as is necessarily 
the case whenever there is a potential concern about tipping point dynamics --
simpler models therefore fit just as well and tend to be selected more often than complex 
ones.  This approach would be appropriate when the dynamics can be expected 
to remain in the region of the training data; for instance, if we only 
considered the forecasting accuracy of the unfished population dynamics under
each model.

@@

### GPDP population dynamics capture larger uncertainty in regions where the data are poor

The parametric models perform worst when they propose a management strategy
outside the range of the observed data. The non-parametric Bayesian approach, 
in contrast, allows a predictive model that expresses a great deal of uncertainty
about the probable dynamics outside the observed range, while retaining very
good predictive accuracy in the range observed.  The management policy 
dictated by the GP balances this uncertainty against
the immediate value of the harvest, and act to stabilize the population 
dynamics in a region of state space in which the predictions can be 
reliably reflected by the data.

@@

Lastly, it should be noted that outside the data, the GP reverts to
the prior, and consequently the choice of the prior can also play a
significant role in determining the optimal policy inferred by the
SDP.  In the examples shown here we have selected a prior that is both
relatively uninformative (due to the broad priors placed on its parameters
$\ell$ and $\sigma$ and simple (mean zero, radial basis function kernel).
In practice, both the choice of mean and the covariance
function may be chosen to confer particular biological properties,
as well as more biologically informed priors for $\ell$ and $\sigma$.
In principle, this may allow a manager to improve the performance of the
GPDP by adding only enough additional detail as is justified.
For instance, it would be possible to use a linear or a Ricker-shaped
mean in the prior without making the much stronger assumption that the
Ricker is the structurally correct model [@Sugeno2013a]. Future research
should focus on identifying criteria to ensure the prior and the value function
are chosen appropriately for the problem at hand.

@@

\appendix
\renewcommand*{\thefigure}{S\arabic{figure}}
\renewcommand*{\thetable}{S\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}

@@

set.A = c(r = p1[1], K = p2[1], theta = p3[1])
set.B = c(r = p1[2], K = p2[2], theta = p3[2])
set.C = c(r = p1[3], K = p2[3], C = p3[3])
set.D = c(r = p1[4], K = p2[4], C = p3[4])
AllenParameterSets <- rbind(set.A, set.B)
MyersParameterSets <- rbind(set.C, set.D)
